---
title: "Hyacinth invasion and control analysis"
author: "DJC"
date: "9/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Analysis for Dan's Chapter 3

Dan conducted a mesocosm experiment that established snail populations in aquatic communities that were invaded (or not) by water hyacinth and then implemented plant management (destruction or removal) after sampling on week seven post establishment. Communities were sampled for 17 weeks from establishment, measuring algal production, snail abundance, snail population total biomass, snail egg laying, and total cercarial production from the entire population.

We're going to use GAMMs to model the "baseline" dynamics and then use additional smooths to account for the interventions. This stems from specific hypothesized contrasts between treatments (e.g., invaded vs. uninvaded; unmanaged invasion vs. the two management options).

We also just summed up cercs to test for differences across the entire experiment, which ignores timing.

```{r echo=FALSE, messages=FALSE}
library(dplyr)
library(ggplot2)
library(cowplot)
theme_set(theme_cowplot())
library(mgcv)
library(glmmTMB)
library(itsadug)
library(emmeans)
```



# Data
First, we need to bring in the data, which are in three sheets (along with a 4th sheet that contains the treatment labels for tanks)

## Egg data
```{r egg_data}
# Egg count master sheet 
eggs = read.csv("C:/RData/Dan dissertation/Summer2019Eggs.csv")

# Egg master sheet has the number of observed eggs (not egg masses!) in each tank that had any. So, we need to sum up and
# add in zeros for those tank-dates where no eggs were found
Eggs = aggregate(Eggs.mass ~ Tank*Week, FUN=sum, data=eggs, drop=F)
Eggs[which(is.na(Eggs[,"Eggs.mass"])),"Eggs.mass"] = 0

# Other data sheets have eggs on "week zero", we want to add NAs so that the data sheets eventually align
Eggs0 = data.frame("Tank" = c(1:10, 12:23, 27:40), "Week" = 0, "Eggs.mass" = NA)
# Egg data is now in correct orientation
Eggs = rbind(Eggs0, Eggs)

```
## Cerc, plant, and Periphyton data

```{r}
# Some of the data is stored in a tank-level master sheet
CercsPlantsPeri = subset(read.csv("Summer2019HyacinthExpMaster.csv"), !Tank %in% c(11, 24:26), select=c("Tank","Week", "Plant.mass", "Cercs", "Peri"))
```


## Snail data 
```{r snail_data}
hya = read.csv("C:/RData/Hyacinth_control_MASTER_041421.csv")
hya[,"Tank"] = as.factor(hya[,"Tank"])
hya[,"Biomass"] = 0.0096*hya[,"Length"]^3

abundance = aggregate(Length ~ Tank*Week, data=hya, FUN=length, drop=F)

labels = subset(read.csv("Dan_hyacinth_labels.csv"), !(Tank %in% c(11, 24:26)))

abundance = cbind(labels,"Week" = abundance$Week, "Abundance" = abundance$Length)
biomass = aggregate(Biomass ~ Tank*Week, data=hya, FUN=sum, drop=F)

tank_summary = cbind(abundance, "Biomass" =  biomass$Biomass, "Eggs" = Eggs$Eggs.mass, "Plants" = CercsPlantsPeri$Plant.mass, "Cercariae" = CercsPlantsPeri$Cercs, "Periphyton" = CercsPlantsPeri$Peri)

# treatment_summary
SEM = function(x){sd(x)/sqrt(na.omit(length(x)))}

# Data from the snail master sheet
mean_abundance = aggregate(Abundance ~ Plant*Management*Week, FUN=mean, data=tank_summary, na.action = na.omit)
SE_abundance = aggregate(Abundance ~ Plant*Management*Week, FUN=SEM, data=tank_summary, na.action = na.omit)

mean_biomass = aggregate(Biomass ~ Plant*Management*Week, FUN=mean, data=tank_summary, na.action = na.omit)
SE_biomass = aggregate(Biomass ~ Plant*Management*Week, FUN=SEM, data=tank_summary, na.action = na.omit)

# Data from egg sheet
mean_eggs = aggregate(Eggs ~ Plant*Management*Week, FUN=mean, data=tank_summary, na.action = na.omit)
SE_eggs = aggregate(Eggs ~ Plant*Management*Week, FUN=SEM, data=tank_summary, na.action = na.omit)
colnames(SE_eggs)[4] = "Eggs_SE"

treatment_summary = data.frame(mean_abundance, "Abundance_SE" = SE_abundance$Abundance,
                               "Biomass" = mean_biomass$Biomass, "Biomass_SE" = SE_biomass$Biomass)

# Data that were collected in the other datasheet
mean_cercs = aggregate(Cercariae ~ Plant*Management*Week, FUN=mean, data=tank_summary, na.action = na.omit)
SE_cercs = aggregate(Cercariae ~ Plant*Management*Week, FUN=SEM, data=tank_summary, na.action = na.omit)
colnames(SE_cercs)[4] = "Cercariae_SE"



mean_peri = aggregate(Periphyton ~ Plant*Management*Week, FUN=mean, data=tank_summary)
SE_peri = aggregate(Periphyton ~ Plant*Management*Week, FUN=SEM, data=tank_summary)
colnames(SE_peri)[4] = "Periphyton_SE"

mean_plants = aggregate(Plants ~ Plant*Management*Week, FUN=mean, data=tank_summary)
SE_plants = aggregate(Plants ~ Plant*Management*Week, FUN=SEM, data=tank_summary)
colnames(SE_plants)[4] = "Plant_SE"

treatment_summary = left_join(treatment_summary, mean_cercs, by=c("Plant", "Management", "Week"))
treatment_summary = left_join(treatment_summary, SE_cercs, by=c("Plant", "Management", "Week"))

treatment_summary = left_join(treatment_summary, mean_peri, by=c("Plant", "Management", "Week"))
treatment_summary = left_join(treatment_summary, SE_peri, by=c("Plant", "Management", "Week"))

treatment_summary = left_join(treatment_summary, mean_plants, by=c("Plant", "Management", "Week"))
treatment_summary = left_join(treatment_summary, SE_plants, by=c("Plant", "Management", "Week"))

treatment_summary = left_join(treatment_summary, mean_eggs, by=c("Plant", "Management", "Week"))
treatment_summary = left_join(treatment_summary, SE_eggs, by=c("Plant", "Management", "Week"))


treatment_summary
```

## The first set of tests correspond to the prediction that hyacinth invasion will alter snail abundance, biomass, periphyton, egg laying, and cercariae

```{r invasion_effects}
invasion = subset(tank_summary, Management == "none ")
invasion[,"IsHyacinth"] = ifelse(invasion[,"Plant"]=="Hyacinth", 1, 0)
invasion[,"Tank"] = as.factor(invasion[,"Tank"])
invasion1 = subset(invasion, is.finite(Abundance))

m1 = gamm(Abundance ~ s(Week)+ s(Week, by = IsHyacinth)+ s(Tank, bs="re"), niterPQL=50,
           family=quasipoisson, correlation=corCAR1(form=~Week|Tank), data=invasion1)
summary(m1$gam)
#plot(m1$gam) # Significantly more snails in Week ~5, significantly fewer in weeks 8+

invasion2 = subset(invasion, is.finite(Biomass))

m2 = gamm(Biomass ~ s(Week)+ s(Week, by = IsHyacinth)+ s(Tank, bs="re"), niterPQL=50,
           family=Gamma(link="log"), correlation=corCAR1(form=~Week|Tank), data=invasion2)
summary(m2$gam)
#plot(m2$gam) # Significantly more biomass in Week ~5, significantly less in weeks 8+

invasion3 = subset(invasion, is.finite(Periphyton))

m3 = gamm(Periphyton ~ s(Week, k=8)+ s(Week, by = IsHyacinth, k=8)+ s(Tank, bs="re"), niterPQL=50,
           family=Gamma(link="log"), correlation=corCAR1(form=~Week|Tank), data=invasion3)
summary(m3$gam)
#plot(m3$gam) # Significantly less peruphyton the whole time

invasion4 = subset(invasion, is.finite(Eggs))

m4 = gamm(Eggs ~ s(Week)+ s(Week, by = IsHyacinth)+ s(Tank, bs="re"), niterPQL=50,
           family=quasipoisson, correlation=corCAR1(form=~Week|Tank), data=invasion4)
summary(m4$gam)
#plot(m4$gam) # Significantly fewer eggs from Week ~2 on

invasion5 = subset(invasion, is.finite(Cercariae))

m5 = gamm(Cercariae ~ s(Week)+ s(Week, by = IsHyacinth)+ s(Tank, bs="re"), niterPQL=50,
           family=quasipoisson, correlation=corCAR1(form=~Week|Tank), data=invasion5)
summary(m5$gam)
#plot(m5$gam) # significantly more cercariae throughout almost the whole experiment

```


```{r fig2}


# make treatment level summary data frame for invasion contrast
invasion_summary = subset(treatment_summary, Management == "none ")
invasion_summary[,"IsHyacinth"] = as.factor(ifelse(invasion_summary[,"Plant"] == "Hyacinth", 1, 0))

# Panel 2A - Periphyton
p2Afit = plot_smooth(m3$gam, view="Week", plot_all=c("IsHyacinth"), rug=F, transform=exp, se=1, shade=T, 
                    ylab="Response", rm.ranef=T, xlab="Time (weeks)", hide.label=T)$fv
p2Afit[,"IsHyacinth"] = as.factor(p2Afit[,"IsHyacinth"])


p2A =  ggplot(data=p2Afit, aes(x=Week, y=fit, group=IsHyacinth, colour=IsHyacinth)) +
  theme(legend.position = "none",
        axis.ticks.length = unit(-1.5, "mm"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size=8, margin = margin(t=3)),
        axis.text.y = element_text(size=8, margin = margin(r=3))) + 
  geom_line() + geom_ribbon(aes(ymin=ll, ymax=ul, fill=IsHyacinth, colour=NA), alpha=0.2) +
  scale_color_manual(values=c("black", "green3"), na.value=NA) +
  scale_fill_manual(values=c("black", "green3"), na.value=NA) + 
  geom_point(data=invasion_summary, aes(x=Week, y=Periphyton, colour=IsHyacinth, group=IsHyacinth)) +
  geom_linerange(data=invasion_summary, aes(x=Week, ymin=Periphyton - Periphyton_SE,
                                        ymax=Periphyton + Periphyton_SE, colour=IsHyacinth), inherit.aes = F) 
p2A


# Panel 2B - Snail abundance
p2Bfit = plot_smooth(m1$gam, view="Week", plot_all=c("IsHyacinth"), rug=F, transform=exp, se=1, shade=T, 
                    ylab="Response", rm.ranef=T, xlab="Time (weeks)", hide.label=T)$fv
p2Bfit[,"IsHyacinth"] = as.factor(p2Bfit[,"IsHyacinth"])


p2B =  ggplot(data=p2Bfit, aes(x=Week, y=fit, group=IsHyacinth, colour=IsHyacinth)) +
  theme(legend.position = "none",
        axis.ticks.length = unit(-1.5, "mm"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size=8, margin = margin(t=3)),
        axis.text.y = element_text(size=8, margin = margin(r=3))) + 
  geom_line() + geom_ribbon(aes(ymin=ll, ymax=ul, fill=IsHyacinth, colour=NA), alpha=0.2) +
  scale_color_manual(values=c("black", "green3"), na.value=NA) +
  scale_fill_manual(values=c("black", "green3"), na.value=NA) + 
  geom_point(data=invasion_summary, aes(x=Week, y=Abundance, colour=IsHyacinth, group=IsHyacinth)) +
  geom_linerange(data=invasion_summary, aes(x=Week, ymin=Abundance - Abundance_SE,
                                        ymax=Abundance + Abundance_SE, colour=IsHyacinth), inherit.aes = F) 
p2B

# Panel 2C - Snail biomass
p2Cfit = plot_smooth(m2$gam, view="Week", plot_all=c("IsHyacinth"), rug=F, transform=exp, se=1, shade=T, 
                    ylab="Response", rm.ranef=T, xlab="Time (weeks)", hide.label=T)$fv
p2Cfit[,"IsHyacinth"] = as.factor(p2Cfit[,"IsHyacinth"])


p2C =  ggplot(data=p2Cfit, aes(x=Week, y=fit, group=IsHyacinth, colour=IsHyacinth)) +
  theme(legend.position = "none",
        axis.ticks.length = unit(-1.5, "mm"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size=8, margin = margin(t=3)),
        axis.text.y = element_text(size=8, margin = margin(r=3))) + 
  geom_line() + geom_ribbon(aes(ymin=ll, ymax=ul, fill=IsHyacinth, colour=NA), alpha=0.2) +
  scale_color_manual(values=c("black", "green3"), na.value=NA) +
  scale_fill_manual(values=c("black", "green3"), na.value=NA) + 
  geom_point(data=invasion_summary, aes(x=Week, y=Biomass, colour=IsHyacinth, group=IsHyacinth)) +
  geom_linerange(data=invasion_summary, aes(x=Week, ymin=Biomass - Biomass_SE,
                                        ymax=Biomass + Biomass_SE, colour=IsHyacinth), inherit.aes = F) 
p2C

# Panel 2D - Snail egg laying
p2Dfit = plot_smooth(m4$gam, view="Week", plot_all=c("IsHyacinth"), rug=F, transform=exp, se=1, shade=T, 
                    ylab="Response", rm.ranef=T, xlab="Time (weeks)", hide.label=T)$fv
p2Dfit[,"IsHyacinth"] = as.factor(p2Dfit[,"IsHyacinth"])


p2D =  ggplot(data=p2Dfit, aes(x=Week, y=fit, group=IsHyacinth, colour=IsHyacinth)) +
  theme(legend.position = "none",
        axis.ticks.length = unit(-1.5, "mm"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size=8, margin = margin(t=3)),
        axis.text.y = element_text(size=8, margin = margin(r=3))) + 
  geom_line() + geom_ribbon(aes(ymin=ll, ymax=ul, fill=IsHyacinth, colour=NA), alpha=0.2) +
  scale_color_manual(values=c("black", "green3"), na.value=NA) +
  scale_fill_manual(values=c("black", "green3"), na.value=NA) + 
  geom_point(data=invasion_summary, aes(x=Week, y=Eggs, colour=IsHyacinth, group=IsHyacinth)) +
  geom_linerange(data=invasion_summary, aes(x=Week, ymin=Eggs - Eggs_SE,
                                        ymax=Eggs + Eggs_SE, colour=IsHyacinth), inherit.aes = F) 
p2D



# Panel 1E - Cecariae
p2Efit = plot_smooth(m5$gam, view="Week", plot_all=c("IsHyacinth"), rug=F, transform=exp, se=1, shade=T, 
                    ylab="Response", rm.ranef=T, xlab="Time (weeks)", hide.label=T)$fv
p2Efit[,"IsHyacinth"] = as.factor(p2Efit[,"IsHyacinth"])


p2E =  ggplot(data=p2Efit, aes(x=Week, y=fit, group=IsHyacinth, colour=IsHyacinth)) +
  theme(legend.position = "none",
        axis.ticks.length = unit(-1.5, "mm"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size=8, margin = margin(t=3)),
        axis.text.y = element_text(size=8, margin = margin(r=3))) + 
  geom_line() + geom_ribbon(aes(ymin=ll, ymax=ul, fill=IsHyacinth, colour=NA), alpha=0.2) +
  scale_color_manual(values=c("black", "green3"), na.value=NA) +
  scale_fill_manual(values=c("black", "green3"), na.value=NA) + 
  geom_point(data=invasion_summary, aes(x=Week, y=Cercariae, colour=IsHyacinth, group=IsHyacinth)) +
  geom_linerange(data=invasion_summary, aes(x=Week, ymin=Cercariae - Cercariae_SE,
                                        ymax=Cercariae + Cercariae_SE, colour=IsHyacinth), inherit.aes = F) 
p2E

spacer = ggplot(data=p2Efit, aes(x=Week, y=fit)) +
  geom_blank() + theme_void()

Fig2 = plot_grid(spacer,p2A, spacer, p2D,
                 spacer,p2B, spacer,p2E,
                 spacer,p2C, spacer, spacer,
                 spacer,spacer, spacer, spacer,
                 align="v", ncol=4, nrow=4, axis="rltb", rel_heights=c(1, 1, 1, 0.1), rel_widths = c(0.08, 1, 0.12, 1)) +
  # y-axis labels
  draw_label("Periphyton productivity ± SE", x=0.03, y=0.84, angle=90, size=10) +
  draw_label("Snail abundance ± SE", x=0.03, y=0.52, angle=90, size=10) +
  draw_label("Snail biomass, mg ± SE", x=0.03, y=0.21, angle=90, size=10) +
  draw_label("Snail eggs ± SE", x=0.54, y=0.84, angle=90, size=10) +
  draw_label("Cercariael production ± SE", x=0.54, y=0.52, angle=90, size=10) +

  # x-axis labels
  draw_label("Week", x = 0.55, y = 0.01, size=10, vjust=0) +
  # panel labels
  draw_label("A", x = 0.16, y = 0.97, size=10) + 
  draw_label("D", x = 0.66, y = 0.97, size=10) + 
  draw_label("B", x = 0.16, y = 0.64, size=10) +
  draw_label("E", x = 0.66, y = 0.64, size=10) +
  draw_label("C", x = 0.16, y = 0.33, size=10) +

  # Legend panel
# legend
  draw_label("Invasion status", x=0.78, y=0.23, size=11)+
  draw_label("•", x=0.71, y=0.20, size=28, colour="black")+ draw_label("Control", x=0.73, y=0.20, size=10, hjust=0)+ # Alt+0149 for bullet point
  draw_label("•", x=0.71, y=0.18, size=28, colour="green3")+ draw_label("Hyacinth", x=0.73, y=0.18, size=10, hjust=0)

Fig2

save_plot("C:/RData/Fig2_Hyacinth.png", Fig2, ncol=2, nrow=3, base_height=2, base_asp = 1.1, dpi=600, units="in")
```



## The second set of tests correspond to the prediction that given hyacinth invasion, management interventions will alter snail abundance, biomass, periphyton, egg laying, and cercariae

```{r management_effects}
management = subset(tank_summary, Plant == "Hyacinth")
head(management)
# "intervened" accounts for a "step effect" of intervention, but have to dummy label by treatment
management[,"intervened"] = ifelse(management[,"Management"]  != "none " & management[,"Week"] > 7, 1, 0)
management[,"removed"] = ifelse(management[,"intervened"] & management[,"Management"] == "remove", 1, 0)
management[,"chopped"] = ifelse(management[,"intervened"] & management[,"Management"] == "Chop", 1, 0)


# "XXX_time" accounts for nonlinear effect of that intervention
management[,"chop_time"] = pmax(0, (management[,"chopped"]*management[,"Week"]) - 7)
management[,"removed_time"] = pmax(0, (management[,"removed"]*management[,"Week"]) - 7)

management[,"Management"] = as.factor(management[,"Management"])
management[,"Tank"] = as.factor(management[,"Tank"])


management1 = subset(management, is.finite(Abundance))

m1 = gamm(Abundance ~ s(Week) + s(chop_time) + s(removed_time) + s(Tank, bs="re"), 
          niterPQL=50, family=quasipoisson, correlation=corCAR1(form=~Week|Tank), data=management1,   
          control=list(msMaxIter=100))
summary(m1$gam)
plot(m1$gam) # No significant diffs

management2 = subset(management, is.finite(Biomass))

m2 = gamm(Biomass ~  s(Week) + s(chop_time) + s(removed_time) + s(Tank, bs="re"), niterPQL=50,
           family=Gamma(link="log"), correlation=corCAR1(form=~Week|Tank), data=management2)
summary(m2$gam)
plot(m2$gam) # Chopped caused a bump in biomass

management3 = subset(management, is.finite(Periphyton))

m3 = gamm(Periphyton ~ s(Week, k=8) + s(chop_time, k=4) + s(removed_time, k=4) + 
          s(Tank, bs="re"), niterPQL=50, family=Gamma(link="log"), correlation=corCAR1(form=~Week|Tank), 
          data=management3)
summary(m3$gam)
plot(m3$gam) # Chopped caused more periphyton over remaining course of experiment

management4 = subset(management, is.finite(Eggs))

m4 = gamm(Eggs ~  s(Week) + s(chop_time) + s(removed_time) + s(Tank, bs="re"), niterPQL=50,
           family=quasipoisson, correlation=corCAR1(form=~Week|Tank), data=management4)
summary(m4$gam)
plot(m4$gam) # Both interventions very significant

management5 = subset(management, is.finite(Cercariae))

m5 = gamm(Cercariae ~ s(Week) + s(chop_time) + s(removed_time) + s(Tank, bs="re"), niterPQL=50,
           family=quasipoisson, correlation=corCAR1(form=~Week|Tank), data=management5)
summary(m5$gam)
plot(m5$gam) # No differences in cercs

```

```{r fig3}
# Data frame to extract predictions
manage_pred_data = subset(management, Tank %in% c(2, 3, 17), select = c("Tank", "Treat", "Week", "chop_time", "removed_time"))



# Panel 3A - periphyton
pred <- predict(m3$gam, newdata=manage_pred_data, type = 'link', se.fit = TRUE, exclude="s(Tank)")
pred$ll = exp(pred$fit - pred$se.fit)
pred$ul = exp(pred$fit + pred$se.fit)
pred$fit = exp(pred$fit)

manage_peri_data = data.frame(manage_pred_data, pred)

management_summary = subset(treatment_summary, Plant == "Hyacinth")
management_summary[,"Treat"] = rep(c("Hyacinth.Chop", "Hyacinth", "Hyacinth.remove"), times=18)


p3A =  ggplot(data=manage_peri_data, aes(x=Week, y=fit, group=Treat, colour=Treat)) +
  theme(legend.position = "none",
        axis.ticks.length = unit(-1.5, "mm"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size=8, margin = margin(t=3)),
        axis.text.y = element_text(size=8, margin = margin(r=3))) + 
        geom_line() + geom_ribbon(data=manage_peri_data, aes(ymin=ll, ymax=ul, fill=Treat, colour=NA), 
            alpha=0.2) +
  geom_point(data=management_summary, aes(x=Week, y=Periphyton, colour=Treat), inherit.aes=F) +
  geom_linerange(data=management_summary, aes(x=Week, ymin=Periphyton - Periphyton_SE,
            ymax=Periphyton + Periphyton_SE, colour=Treat), inherit.aes = F) +
  scale_color_manual(values=c("green3", "black", "blue"), na.value=NA) +
  scale_fill_manual(values=c("green3", "black", "blue"), na.value=NA)

p3A

# Panel 3B - Snail abundance
pred <- predict(m1$gam, newdata=manage_pred_data, type = 'link', se.fit = TRUE, exclude="s(Tank)")
pred$ll = exp(pred$fit - pred$se.fit)
pred$ul = exp(pred$fit + pred$se.fit)
pred$fit = exp(pred$fit)

manage_abund_data = data.frame(manage_pred_data, pred)


p3B =  ggplot(data=manage_abund_data, aes(x=Week, y=fit, group=Treat, colour=Treat)) +
  theme(legend.position = "none",
        axis.ticks.length = unit(-1.5, "mm"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size=8, margin = margin(t=3)),
        axis.text.y = element_text(size=8, margin = margin(r=3))) + 
        geom_line() + geom_ribbon(data=manage_abund_data, aes(ymin=ll, ymax=ul, fill=Treat, colour=NA), 
            alpha=0.2) +
  geom_point(data=management_summary, aes(x=Week, y=Abundance, colour=Treat), inherit.aes=F) +
  geom_linerange(data=management_summary, aes(x=Week, ymin=Abundance - Abundance_SE,
            ymax=Abundance + Abundance_SE, colour=Treat), inherit.aes = F) +
  scale_color_manual(values=c("green3", "black", "blue"), na.value=NA) +
  scale_fill_manual(values=c("green3", "black", "blue"), na.value=NA)

p3B

# Panel 3C - snail biomass
pred <- predict(m2$gam, newdata=manage_pred_data, type = 'link', se.fit = TRUE, exclude="s(Tank)")
pred$ll = exp(pred$fit - pred$se.fit)
pred$ul = exp(pred$fit + pred$se.fit)
pred$fit = exp(pred$fit)

manage_biomass_data = data.frame(manage_pred_data, pred)

p3C =  ggplot(data=manage_biomass_data, aes(x=Week, y=fit, group=Treat, colour=Treat)) +
  theme(legend.position = "none",
        axis.ticks.length = unit(-1.5, "mm"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size=8, margin = margin(t=3)),
        axis.text.y = element_text(size=8, margin = margin(r=3))) + 
        geom_line() + geom_ribbon(data=manage_biomass_data, aes(ymin=ll, ymax=ul, fill=Treat, colour=NA), 
            alpha=0.2) +
  geom_point(data=management_summary, aes(x=Week, y=Biomass, colour=Treat), inherit.aes=F) +
  geom_linerange(data=management_summary, aes(x=Week, ymin=Biomass - Biomass_SE,
            ymax=Biomass + Biomass_SE, colour=Treat), inherit.aes = F) +
  scale_color_manual(values=c("green3", "black", "blue"), na.value=NA) +
  scale_fill_manual(values=c("green3", "black", "blue"), na.value=NA)

p3C


# Panel 3D - snail eggs
pred <- predict(m4$gam, newdata=manage_pred_data, type = 'link', se.fit = TRUE, exclude="s(Tank)")
pred$ll = exp(pred$fit - pred$se.fit)
pred$ul = exp(pred$fit + pred$se.fit)
pred$fit = exp(pred$fit)

manage_egg_data = data.frame(manage_pred_data, pred)

p3D =  ggplot(data=manage_egg_data, aes(x=Week, y=fit, group=Treat, colour=Treat)) +
  theme(legend.position = "none",
        axis.ticks.length = unit(-1.5, "mm"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size=8, margin = margin(t=3)),
        axis.text.y = element_text(size=8, margin = margin(r=3))) + 
        geom_line() + geom_ribbon(data=manage_egg_data, aes(ymin=ll, ymax=ul, fill=Treat, colour=NA), 
            alpha=0.2) +
  geom_point(data=management_summary, aes(x=Week, y=Eggs, colour=Treat), inherit.aes=F) +
  geom_linerange(data=management_summary, aes(x=Week, ymin=Eggs - Eggs_SE,
            ymax=Eggs + Eggs_SE, colour=Treat), inherit.aes = F) +
  scale_color_manual(values=c("green3", "black", "blue"), na.value=NA) +
  scale_fill_manual(values=c("green3", "black", "blue"), na.value=NA)

p3D




# Panel 3E - Cecariae
manage_pred_data = subset(manage_pred_data, Week >= 4)
pred <- predict(m5$gam, newdata=manage_pred_data, type = 'link', se.fit = TRUE, exclude="s(Tank)")
pred$ll = exp(pred$fit - pred$se.fit)
pred$ul = exp(pred$fit + pred$se.fit)
pred$fit = exp(pred$fit)

manage_cerc_data = data.frame(manage_pred_data, pred)

p3E =  ggplot(data=manage_cerc_data, aes(x=Week, y=fit, group=Treat, colour=Treat)) +
  theme(legend.position = "none",
        axis.ticks.length = unit(-1.5, "mm"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size=8, margin = margin(t=3)),
        axis.text.y = element_text(size=8, margin = margin(r=3))) + 
        geom_line() + geom_ribbon(data=manage_cerc_data, aes(ymin=ll, ymax=ul, fill=Treat, colour=NA), 
            alpha=0.2) +
  geom_point(data=management_summary, aes(x=Week, y=Cercariae, colour=Treat), inherit.aes=F) +
  geom_linerange(data=management_summary, aes(x=Week, ymin=Cercariae - Cercariae_SE,
            ymax=Cercariae + Cercariae_SE, colour=Treat), inherit.aes = F) +
  scale_color_manual(values=c("green3", "black", "blue"), na.value=NA) +
  scale_fill_manual(values=c("green3", "black", "blue"), na.value=NA)

p3E

spacer = ggplot(data=p1Efit, aes(x=Week, y=fit)) +
  geom_blank() + theme_void()

Fig3 = plot_grid(spacer,p3A, spacer, p3D,
                 spacer,p3B, spacer,p3E,
                 spacer,p3C, spacer, spacer,
                 spacer,spacer, spacer, spacer,
                 align="v", ncol=4, nrow=4, axis="rltb", rel_heights=c(1, 1, 1, 0.1), rel_widths = c(0.08, 1, 0.12, 1)) +
  # y-axis labels
  draw_label("Periphyton productivity ± SE", x=0.03, y=0.84, angle=90, size=10) +
  draw_label("Snail abundance ± SE", x=0.03, y=0.52, angle=90, size=10) +
  draw_label("Snail biomass, mg ± SE", x=0.03, y=0.21, angle=90, size=10) +
  draw_label("Snail eggs ± SE", x=0.54, y=0.84, angle=90, size=10) +
  draw_label("Cercariael production ± SE", x=0.54, y=0.52, angle=90, size=10) +

  
  # x-axis labels
  draw_label("Week", x = 0.55, y = 0.01, size=10, vjust=0) +
  # panel labels
  draw_label("A", x = 0.16, y = 0.97, size=10) + 
  draw_label("D", x = 0.66, y = 0.97, size=10) + 
  draw_label("B", x = 0.16, y = 0.64, size=10) +
  draw_label("E", x = 0.66, y = 0.64, size=10) +
  draw_label("C", x = 0.16, y = 0.33, size=10) +

  # Legend panel
# legend
  draw_label("Management status", x=0.78, y=0.23, size=11)+
  draw_label("•", x=0.71, y=0.20, size=28, colour="green3")+ draw_label("Unmanaged", x=0.73, y=0.20, size=10, hjust=0)+ # Alt+0149 for bullet point
  draw_label("•", x=0.71, y=0.18, size=28, colour="black")+ draw_label("Destruction", x=0.73, y=0.18, size=10, hjust=0)+
    draw_label("•", x=0.71, y=0.16, size=28, colour="blue")+ draw_label("Removal", x=0.73, y=0.16, size=10, hjust=0)

Fig3

save_plot("C:/RData/Fig3_Hyacinth.png", Fig3, ncol=2, nrow=3, base_height=2, base_asp = 1.1, dpi=600, units="in")


```




## The third test is a direct analysis of cumulative cercarial production across all of the treatments compared against the control
```{r}
cumulative_cercs = aggregate(Cercariae ~ Tank*Plant*Management, FUN=sum, data=subset(tank_summary, Week > 0))

# drop control-chop
#cumulative_cercs = subset(cumulative_cercs, Plant != "Control" | Management != "Chop")
cumulative_cercs[,"Treatment"] = paste(cumulative_cercs[,"Plant"], cumulative_cercs[,"Management"])
# Rename for plotting
cumulative_cercs[,"Treatment"][which(cumulative_cercs[,"Treatment"] == "Control none ")] = "Uninvaded\ncontrol"
cumulative_cercs[,"Treatment"][which(cumulative_cercs[,"Treatment"] == "Control Chop")] = "Uninvaded control\nplus detritus"
cumulative_cercs[,"Treatment"][which(cumulative_cercs[,"Treatment"] == "Hyacinth none ")] = "Unmanaged\nhyacinth"
cumulative_cercs[,"Treatment"][which(cumulative_cercs[,"Treatment"] == "Hyacinth Chop")] = "Hyacinth\ndestruction"
cumulative_cercs[,"Treatment"][which(cumulative_cercs[,"Treatment"] == "Hyacinth remove")] = "Hyacinth\nremoval"
cumulative_cercs[,"Treatment"] <- factor(cumulative_cercs[,"Treatment"], levels = c("Uninvaded\ncontrol", "Uninvaded control\nplus detritus", "Unmanaged\nhyacinth", "Hyacinth\ndestruction", "Hyacinth\nremoval"))



m1 = glmmTMB(Cercariae ~ Treatment, data=cumulative_cercs, family="nbinom2")
summary(m1)

Fig4 = ggplot(data = cumulative_cercs, aes(x = Treatment, y = Cercariae)) + geom_boxplot()  +
      theme(axis.ticks.length = unit(-1.5, "mm"),
            axis.title = element_text(size=14),
            axis.text = element_text(size=11)) +  xlab("Treatment") +  ylab("Cumulative Cercarial production") +
      annotate("text", x = 3, y = 4500, label = "*", size=10) + annotate("text", x = 4, y = 5500, label = "*", size=10) 

Fig4
ggsave(filename = "C:/RData/Fig4_Hyacinth.png", plot=Fig4, scale=1)

```

